---
title: "BDS Toolbox: DV Assignment 1 Part 2"
author: "Group X; Names: "
date: "`r Sys.Date()`"
output: html_document
---

## Setup environment
```{r setup environment, include = TRUE, eval=FALSE}
knitr::opts_chunk$set(echo = TRUE) # set default echo = TRUE for all code blocks
rm(list = ls()) # Remove any existing objects in memory
```

## A. Research question

Specify your research question here. 
It is also a good idea to take notes from your discussion of the research question and add those here.  

## B. Connect to Git and load data

State the name of your github repository here. If you encountered issues when connecting to the server, or cloning the repository, let us know. 

```{r requirements}
# load packages, set up database connection
```

```{r load-data}
# code to load data for your final visualization 
```

It is a good idea to take notes from your discussion about the data, and add those here. 


```{r data-prep}
# code to clean data for your final visualization 
```

## C. First visualization 

Include code and a description of your first visualization here. If you have specific things you would like feedback on, you can indicate that here too. 

```{r plot-1}
# code to produce your first visualization
```

**We will only grade the visualization pertaining to part C. If you make exploratory visualizations in part B, feel free to include them in that section, as it may help us to understand your process. If you make more than one visualization in part C, please indicate clearly which visualization we should look at when grading**. 
